When does vapor pressure deficit drive or reduce evapotranspiration?\footnote{Primary Research Article}

### 2018-03-01 doc/paper/vpd_et_paper.tex
done with abstract

\title{When does vapor pressure deficit drive or reduce evapotranspiration?}

@@

\begin{keypoints}
\item one
\item two
\end{keypoints}

\begin{abstract}
= enter abstract here =
\end{abstract}

>>>>

\title{When does vapor pressure deficit drive or reduce evapotranspiration?\footnote{Primary Research Article}}

@@

\begin{keypoints}
\item hydrometeorology, evapotranspiration, aridity, vapor pressure deficit, Clausius-Clapeyron relation, plant physiology, stomatal response
\end{keypoints}

\begin{abstract}
  Increases to vapor pressure deficit (VPD) and atmospheric demand for water are expected with rising CO$_2$. While increased evapotranspiration (ET) in response to increased atmospheric demand seems intuitive, plants are capable of reducing ET in response to increased VPD by closing their stomata, in an effot to conserve water. Here we examine which effect dominates response to increasing VPD: atmospheric demand and increases in ET, or plant physiological response (stomata closure) and decreases in ET. We use Penman-Monteith, combined with semi-empirical optimal stomatal regulation theory and underlying water use efficiency, to develop a theoretical framework for understanding how ET responds to increases in VPD.
  The theory suggests that for most environmental conditions and plant types, plant physiological response dominates and ET decreases with increasing VPD. Plants that are evolved or bred to prioritize primary production over water conservation (e.g. crops) exhibit a higher likelihood of atmospheric demand-driven response (ET increasing). However for forest, grass, and shrub plant types, ET more frequently decreases than increases with rising VPD. Therefore, we may expect reductions in ET due to CO$_2$-induced increases in VPD. This work serves as an example of the utility of our simplified framework for disentangling land-atmosphere feedbacks at multiple scales, including the characterization of ET response in an atmospherically drier, enriched CO$_2$ world.

\end{abstract}
